Quasars Probing Quasars IX. Probing the Circumgalactic Flows of Quasars [v1.1]

### 2017-02-19 QPQ9/qpq9_draft.tex
Generate stack figures, wrote draft

From our QPQ survey\footnote{http://www.qpqsurvey.org} (QPQ1), we have analyzed a subset of 195 
systems that pass within 300\,kpc physical from a foreground quasar with $z > 1.6$. We restrict
 the sample to foreground quasars with redshift measured from \ion{Mg}{2}\,2800, 
[\ion{O}{3}]\,5007, H$\alpha$, or H$\beta$ emission, giving a precision of 
$400\,{\rm km\,s^{-1}}$ or lower. 
%#[\ion{O}{3}] has the smallest dispersion of $44\,{\rm km\,s^{-1}}$ about the systemic redshift, 
%and we analyze the sub-sample with [\ion{O}{3}] redshifts separately. 
The [\ion{O}{3}] line has an average blueshift of $27\,{\rm km\,s^{-1}}$ about the systemic 
redshift \citep{Boroson05}, which has been added when we compute the 
redshift of the line. Systemic redshifts measured from \ion{Mg}{2} have a precision of 
$272\,{\rm km\,s^{-1}}$, and we have taken into account the median redshift of 
$97\,{\rm km\,s^{-1}}$ of \ion{Mg}{2} from [\ion{O}{3}] \citep{Richards+02}. In QPQ8. we have 
quantified the precision of H$\alpha$ and H$\beta$ to be 
$300\,{\rm km\,s^{-1}}$ and $392\,{\rm km\,s^{-1}}$ respectively. 
%\citep{HewittWild10}. 
Of the 195 sightlines, 13 have spectral resolution $> 5000$ from echellette or echelle 
observations. This high-dispersion sample has been analyzed separately in QPQ8. 
Figure~\ref{fig:experiment} summarizes the experimental design. We refer the reader to previous 
QPQ publications for the redshift centroiding algorithm, and details on data reduction and 
continuum normalization (QPQ6, QPQ8).

@@

In Table~\ref{tab:summ}, we list the sample size, the median redshift, and the median projected 
separation between the pair of the sub-samples that survive the above selection criteria for 
\ion{C}{2}\,1334, \ion{C}{4}\,1548, and \ion{Mg}{2}\,2796 respectively.

@@

In QPQ8, we measured the velocity widths of the \ion{C}{2}~1334
and \ion{C}{4}~1548 transitions finding that the CGM
frequently exhibits large flows:
median DV{1334},DV{1548} = $XX,XX kms$ with and
RMS DRMS{1334},DRMS{1548} = $XX, XX kms$ from
a sample of ncii\ and nciv\ systems.
These velocity fields exceed all previous measurements from
galaxies and/or absorption systems at any epoch.  For
Figure~\ref{fig:stack}, we present these absorption lines
and \ion{H}{1}~lya, also including lower-dispersion 
sightlines meeting the selection criteria.

To assess the `average' absorption of these transitions,
we have combined the data with a novel technique.
We have fitted the combined dataset using a cubic b-spline
with knots at every 100\,kms.  We weighted each data
point by the velocity width of the pixel to insure that 
each sightline makes an equal contribution to the stack.
These b-spline profiles are presented in the figure
and a trivial integration gives the reported equivalent
widths.
[consider down-weighting systems with $<3\sigma$ EW]
[Comnent on EW from 'standard' technique]

Two results are evident from the figure:
 (i) the data exhibit excess absorptoin at 
 $mdv \approx 0 kms$;
 (ii) the absorption is skewed towards positive velocities
 $mdv > 0 mkms$.
We measure absorption-weigthed ($1-f$) centroids
of $mdv = XX-XX mkms$ for these transitions
(Table~\ref{tab:summ}).
This suggests an asymmetry that contradicts the [standard]
expectation.

Given the small sample size, one must 
scrutinize the statistical significance of the measured offset.
Indeed, a small offset $mdv < 10 kms$ may be expected
even with large samples.  To assess the statistical significance,
we have perfomed a boostrap analysis by randomly resampling
from the full sample. [DISCUSS]

[Discuss systematics here??]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
\label{sec:discussion}

@@

%%%%%%%%%%%%%%%%
\acknowledgements

@@

\bibliographystyle{/u/xavier/paper/Bibli/apj}
\bibliography{/u/xavier/paper/Bibli/allrefs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\input{../Tables/tab_sub_qpq7_spectra.tex}
\input{Tables/tab_summ.tex}
%\input{Tables/tab_summ_Hab.tex}
%\input{tab1_sub.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}
\includegraphics[width=5in]{Figures/fig_experiment.pdf}
%\includegraphics[width=5in,angle=90]{../Figuref1.eps}
\caption{Experimental design
}
\label{fig:exp}
\end{figure}

>>>>

From our QPQ survey\footnote{http://www.qpqsurvey.org} (QPQ1), we have analyzed a subset of  
systems that pass within 300\,kpc physical from a foreground quasar with $z > 1.6$. We restrict
 the sample to foreground quasars with redshift measured from \ion{Mg}{2}\,2800, 
[\ion{O}{3}]\,5007, H$\alpha$, or H$\beta$ emission, giving a precision of 
$400\,{\rm km\,s^{-1}}$ or lower. 
[\ion{O}{3}] emission-line redshifts have the smallest dispersion of $44\,{\rm km\,s^{-1}}$ about 
the systemic redshift, and we analyze the sub-sample with [\ion{O}{3}] redshifts separately. 
The [\ion{O}{3}] line has an average blueshift of $27\,{\rm km\,s^{-1}}$ about the systemic 
redshift \citep{Boroson05}, which has been added when we compute the 
redshift of the line. Systemic redshifts measured from \ion{Mg}{2} have a precision of 
$272\,{\rm km\,s^{-1}}$, and we have taken into account the median redshift of 
$97\,{\rm km\,s^{-1}}$ of \ion{Mg}{2} from [\ion{O}{3}] \citep{Richards+02}. In QPQ8. we have 
quantified the precision of H$\alpha$ and H$\beta$ to be 
$300\,{\rm km\,s^{-1}}$ and $392\,{\rm km\,s^{-1}}$ respectively. We further supplement the 
dataset with quasar pairs selected from the igmspec 
database\footnote{https://github.com/specdb/igmspec}, which comes with the package of software for
accessing spectra of sources useful for probing the intergalactic medium \citep{Prochaska17}. 
We reach a total of 195 sightlines in the final sample. Among them, 13 have spectral resolution 
$> 5000$ from echellette or echelle observations and have been analyzed separately in QPQ8. 
Figure~\ref{fig:experiment} summarizes the experimental design. We refer the reader to previous 
QPQ publications for the redshift centroiding algorithm, and details on data reduction and 
continuum normalization (QPQ6, QPQ8).

@@

In Table~\ref{tab:summary}, we list the sample size, the median redshift, and the median projected 
separation of the quasar pairs that survive the above selection criteria for \ion{C}{2}\,1334, 
\ion{C}{4}\,1548, and \ion{Mg}{2}\,2796 respectively. In Table~\ref{tab:summary_OIII}, we provide 
the summary for the sub-sample with $z_{\rm fg}$ measured from [\ion{O}{3}].

@@

In QPQ8, we measured the velocity widths of the \ion{C}{2}\,1334 and \ion{C}{4}\,1548 transitions, 
finding that the CGM frequently exhibits large flows. From a sample of 7 \ion{C}{2} systems and 10 
\ion{C}{4} systems, we measured the velocity interval that encompasses 90\% of the total optical 
depth, $\Delta v_90$, and the $1\sigma$ dispersion relative to the profile centroid, $\sigma_v$. 
The median $\Delta v_90$ is $555\,{\rm km\,s^{-1}}$ for \ion{C}{2}\,1334 and is 
$342\,{\rm km\,s^{-1}}$ for \ion{C}{4}\,1548. The median $\sigma_v$ is $249\,{\rm km\,s^{-1}}$ for 
\ion{C}{2}\,1334 and is $495\,{\rm km\,s^{-1}}$ for \ion{C}{4}\,1548. These velocity fields exceed 
all previous measurements fromgalaxies and/or absorption systems at any epoch.  
In Figure~\ref{fig:stack_CII}, Figure~\ref{fig:stack_CIV}, and Figure~\ref{fig:stack_MgII}, we 
present mean and median stacks of \ion{C}{2}\,1334, \ion{C}{4}\,1548, and \ion{Mg}{2}2796 
absorption, including lower-dispersion sightlines meeting the selection criteria. For interpreting 
the kinematics, we will focus on the analysis results of the \ion{C}{2} mean stack. The median 
stacks have lower absorption equivalent widths, while \ion{C}{4} and \ion{Mg}{2} are doublet 
transitions and hence interpret their kinematics becomes more challenging.

%To assess the average absorption of these transitions,
%we have combined the data with a novel technique.
%We have fitted the combined dataset using a cubic b-spline
%with knots at every 100\,kms.  We weighted each data
%point by the velocity width of the pixel to insure that 
%each sightline makes an equal contribution to the stack.
%These b-spline profiles are presented in the figure
%and a trivial integration gives the reported equivalent
%widths.
%[consider down-weighting systems with $<3\sigma$ EW]
%[Comnent on EW from 'standard' technique]

Two results are evident from Figure~\ref{fig:stack_II}: (i) the mean stack exhibits excess 
absorption spanning a large velocity width; (ii) the mean absorption is skewed toward positive 
velocities. Our sign convention is that positive velocities are flowing away. To model the 
absorption, we introduce a Gaussian profile while allowing the constant continuum level to vary. 
From the $\chi^2$ best-fit to the data, we measure the centroid to be $+170\,{\rm km\,s^{-1}}$ and 
the $1\sigma$ dispersion to be $388\,{\rm km\,s^{-1}}$. The centroid suggests an asymmetry that 
contradicts the standard expectation, while the dispersion is larger than  The median stack, on 
the other hand, shows weaker absorption, and the Gaussian model has a centroid closer to zero.

We also create mean and median 
stack for the sub-sample with [\ion{O}{3}] redshifts and model the absorption with Gaussian 
best-fit. The mean stack for this sub-sample has a centroid at $+207\,{\rm km\,s^{-1}}$, and a 
dispersion of $355\,{\rm km\,s^{-1}}$. The median stack of this sub-sample has a centroid at 
$60\,{\rm km\,s^{-1}}$ and a dispersion of $307\,{\rm km\,s^{-1}}$. These values are consistent 
with the stacks of full sample.

To model the mean and median stacks of \ion{C}{4}\,1548 and \ion{Mg}{2}\,1796, we introduce two 
Gaussians with separation equal to the doublet separation of $498\,{\rm km\,s^{-1}}$ and 
$769\,{\rm km\,s^{-1}}$ respectively, and tie the dispersion of the two lines in a doublet. The 
modeling results show that the \ion{C}{4} and \ion{Mg}{2} absorption are consistent with 
\ion{C}{2}, that the velocity field is large and skewed toward positive velocities.

The above analysis are summarized in Table~\ref{tab:summary} and 
Table~\ref{tab:summary_OIII}. The Gaussian and continuum models are overplotted on the data stacks 
in Figure~\ref{fig:stack_CII}, Figure~\ref{fig:stack_CIV}, and Figure~\ref{fig:stack_MgII}.

\subsection{Significance of the asymmetry}
\label{sec:sigificance_+ve}

Given the small sample size, one must scrutinize the statistical significance of the measured 
offsets from $0\,{\rm km\,s^{-1}}$. To assess the statistical significance, we perform a boostrap 
analysis by randomly sampling from the full sample 10000 times. We measure the pixel optical 
depth-weighted centroid of each bootstrap realization, and 
quote the standard deviation in the bootstrap realizations to be the uncertainty of the 
centroid of the data. The uncertainties are comparable 
to the measured offsets, indicating large intrinsic variation in quasar CGM environments.

One may also ask whether the measured positive offsets come from systematic biases in redshift 
measurements due to the Baldwin effect \citep{Baldwin77}. \cite{Shen16} reported that, 
the [\ion{O}{3}] emission of $z\sim2$ quasars is more asymmetric and weaker than that in typically 
less luminous low-$z$ quasars. To test for this potential source of systematic bias, we consider 
the mean stack at \ion{C}{2}\,1334. We replace the [\ion{O}{3}] redshifts by a \ion{Mg}{2}, 
H$\alpha$, or H$\beta$ redshift, when available, and create a mean stack for new 
sample. We have a redshift replacement for 13 out of the 14 systems with [\ion{O}{3}] redshifts in 
the original sample. The new stack is similiar in velocity structure and shows a similar positive 
offset. We thus conclude that our algorithm for measuring emission-line redshift is not biased by 
the blue wing of the [\ion{O}{3}] emission-line.

We also generate a stack spectrum for \ion{Mg}{2} for lower redshift quasar pairs, using the same 
selection criteria as the main QPQ9 sample. The quasar pairs are selected from the igmspec 
data base. The 
are 231 quasar pairs selected, with a median $z_{\rm fg}$ of 0.90 and a median $\R_\perp$ of 
208\,kpc. We present the mean and median stack in Figure~\ref{fig:stack_MgII_z1}. The absorption 
is weaker than the $z\approx2$ QPQ9 main sample. Gaussian absorption models of the \ion{Mg}[2} 
doublet fitted to the mean stack recovers a centroid of $47\pm338\,{\rm km\,s^{-1}}$ and a 
dispersion of $119\,{\rm km\,s^{-1}}$, while the fit to the median stack recovers a centroid of 
$14\pm500\,{\rm km\,s^{-1}}$ and a dispersion of $211\,{\rm km\,s^{-1}}$. The offsets 
from $0\,{\rm km\,s^{-1}}$ are smaller than the offsets in the $z\approx2$ sample.

To conclude, the large uncertainties in the velocity centroids represent intrinsic variation not 
redshift errors. Together with the possible difference in $z\approx1$ and $\zapprox2$ quasar 
CGM environments revealed in the stacks for \ion{Mg}{2}, the hint of asymmetry warrants an attempt 
for explanation.

%Indeed, a small offset $mdv < 10 kms$ may be expected even with large samples.

\subsection{Significance of the high-velocity flows}
\label{sec:significance_width}

The dispersion in the \ion{C}{2} mean stack is a result of the intrinsic velocity width of the 
absorption combined with the error in redshift measurements. Using the relation 
$\sigma_v^2 = \sigma_{\rm intrinsic}^2 + \sigma_{error}^2$, we solve for the intrinsic dispersion 
in the stack of the full sample and that in the stack of the sub-sample with $z_{\rm fg}$ measured 
from [\ion{O}{3}]. We recover $\sigma_{\rm intrinsic}^{\rm full}=317\,{\rm km\,s^{-1}}$ and 
$\sigma_{\rm intrinsic}^{\rm [OIII]}=352\,{\rm km\,s^{-1}}$. In QPQ8, we reported the 
line-of-sight velocity dispersion typical of QPQ dark matter halo mass is 
$\sigma_{\rm 1D}=212\,{\rm km\,s^{-1}}$. We investigate whether gravitational motions and Hubble 
flows are sufficient to reproduce the dispersion in the \ion{C}{2} mean stack using Monte Carlo 
methods.

We assume all \ion{C}{2} 
systems arise in optically thick absorbers, and adopt the clustering analysis results of QPQ6. For 
each quasar pair, we calculate the expected number of optically thick absorbers within 
$\pm3000\,{\rm km\,s^{-1}}$ at a distance $R_\perp$ from the foreground quasar and at 
$z_{\rm fg}$. Then we generate 1000 mock spectra. The number of absorbers for each mock sightline 
is randomly selected from a Poisson distribution with mean equals the average number of absorbers. 
The absorbers are randomly assigned Hubble velocities with probabilities according to the 
quasar-absorber correlation functions. The absorbers are randomly assigned additional peculiar 
velocities drawn from a normal distribution with mean equals $0\,{\rm km\,s^{-1}}$ and standard 
deviation equals $\sigma_{\rm 1D}$ of the typical dark matter halo mass. For each absorber, we 
assume a rest equivalent width for \ion{C}{2} and a Gaussian absorption profile. We repeat the 
above procedure for all 34 quasar pairs, and create a mean stack of the 34000 mock spectra 
generated. We fit a Gaussian absorption profile multiplied to a constant continuum level to model 
the stack of mock spectra. We adjust the rest equivalent width adopted for the \ion{C}{2} 
absorbers until the ampltiude of the best-fit Gaussian of the stack of mock spectra matches the 
amplitude of the stack of the observational data. We find that an equivalent width of 0.6\,\AA \ 
well reproduces the amptlitude, and we note that the dispersion in the Gaussian is insensitive to 
the equivalent width or absorption profile for one absorber. The resulting stack of mock spectra 
has a $1\sigma$ dispersion of $\approx240\,{\rm km\,s^{-1}}$, which is less than the 
intrinsic dispersion in the data. In order to recover an intrinsic dispersion of 
$317\text{--}352\,{\rm km\,s^{-1}}$, a peculiar velocity of $\approx280\,{\rm km\,s^{-1}}$ is 
required. We thus conclude the observed velocity width likely requires outflows.

\section{Discussion}
\label{sec:discussion}

@@

\acknowledgements

@@

\bibliographystyle{/lwymarie/Documents/Bibli/apj}
\bibliography{/lwymarie/Documents/Bibli/allrefs}

\input{Tables/tab_summ.tex}
\input{Tables/tab_summ_OIII.tex}

@@

\begin{figure}
\includegraphics[width=3.5in]{Figures/fig_experiment.pdf}
%\includegraphics[width=5in,angle=90]{../Figuref1.eps}
\caption{These panels summarize properties of the QPQ9 dataset. The QPQ survey selects 
quasar pairs of projected separation $R_\perp < 300$\,kpc and $z_{\rm fg}>1.6$. Assuming that the 
foreground quasars emit isotropically and at a distance equal to the impact parameter, we estimate 
enhancement in the UV flux relative to the extragalactic UV background $g_{\rm UV}$. Smaller 
symbols correspond to foreground quasars with the most precise redshift measurement from 
[\ion{O}{3}]\,5007 emission, while larger symbols correspond to foreground quasars with redshift 
measurement from \ion{Mg}{2}\,2796, H$\alpha$, or H$\beta$ emission. The top panel shows 
quasar pairs with coverage of \ion{C}{2}\,1334 at $z_{\rm fg}$ in the background quasar spectra. 
The middle panel shows pairs with coverage of \ion{C}{4}\,1548. The bottom panel shows 
pairs with coverage of \ion{Mg}{2}\,2796. 
}
\label{fig:experiment}
\end{figure}
